---
title: "SNA"
author: "Piombini Davide, Zilli Federico"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '')
```

## Get data
```{r}
collabs <- read.csv('data/FilteredCollaboration.csv')
#Remove id column
collabs <- collabs[,-1]

artists <- read.csv('data/PopularityAndGenres.csv')
artists <- artists[,-1]
dim(artists)
```

## Build Network

```{r, echo=False}
library(igraph)
network <- graph_from_data_frame(collabs, directed = FALSE, vertices = artists)
```

```{r}
print(gsize(network))
print(gorder(network))
```

## Print network
```{r}
# Customize vertex and edge attributes
vertex_color <- "red"  # Color of vertices
vertex_size <- 5      # Size of vertices
edge_width <- 1         # Width of edges

png(filename="Network.png")
plot(network,
     vertex.color = vertex_color,
     vertex.size = vertex_size,
     vertex.label = NA,
     edge.width = edge_width)

dev.off()
```

## Calculate Network values
```{r}
centrality <- degree(network, v = V(network))
artists$num_collab <- centrality[match(artists$artist, V(network)$name)]
```

```{r}
cns <- closeness(network, v = V(network))
artists$closeness <- cns[match(artists$artist, V(network)$name)]
```

```{r}
btw <- betweenness(network, v = V(network))
artists$betweenness <- btw[match(artists$artist, V(network)$name)]
```

## Models

```{r}
model1 <- lm(popularity ~ num_collab, data=artists)

# Print model summary
summary(model1)
```

The output you provided is from a linear regression analysis that aims to explore the relationship between an artist's popularity and the number of collaborations they have on Spotify.

The regression model suggests that there is a statistically significant relationship between an artist's popularity (dependent variable) and the number of collaborations they have (independent variable). The estimated coefficient for the "num_collab" variable is 0.52072, indicating that, on average, each additional collaboration is associated with an increase of approximately 0.52072 units in popularity, holding other factors constant.

To determine the significance of this relationship, the t-value and p-value are provided. The t-value for the "num_collab" coefficient is 7.808, which indicates that the estimated coefficient is significantly different from zero. The corresponding p-value is 6.98e-14, which is extremely small. Since this p-value is below the conventional threshold of 0.05, we can reject the null hypothesis and conclude that the number of collaborations is indeed associated with an artist's popularity.

The multiple R-squared value of 0.1498 indicates that the number of collaborations explains approximately 14.98% of the total variation in artist popularity. This means that other factors not included in the model may also influence an artist's success.

However, it's important to note that correlation does not imply causation. While the analysis shows a positive relationship between collaborations and popularity, it does not necessarily mean that more collaborations directly cause an artist to be successful. There may be other underlying factors or mechanisms that contribute to an artist's success.

Therefore, based on the results of this regression analysis, we have evidence to support Hypothesis 1, suggesting that artists who have collaborated with a larger number of other artists on Spotify are more likely to be successful compared to artists who have collaborated with fewer artists. However, it's important to consider additional factors and conduct further research to gain a comprehensive understanding of the relationship between collaborations and an artist's success.


```{r}
#Create scatterplot with regression line
library(ggplot2)
ggplot(artists, aes(x=num_collab, y=popularity)) +
geom_point() +
geom_smooth(method="lm", se=FALSE)
```
```{r}
plot(model1)
```



```{r}
model2 <- lm(popularity ~ num_genres, data=artists)

# Print model summary
summary(model2)
```
```{r}
#Create scatterplot with regression line
ggplot(artists, aes(x=num_genres, y=popularity)) +
geom_point() +
geom_smooth(method="lm", se=FALSE)
```
The output you provided is from a linear regression analysis, where the dependent variable is the "popularity" of artists and the independent variable is the "num_genres," representing the number of genres an artist is associated with. Here's a brief comment on the output:

The regression model suggests that there is a statistically significant relationship between an artist's popularity and the number of genres they are associated with. The estimated coefficient for the "num_genres" variable is 7.8824, indicating that, on average, each additional genre an artist is associated with is associated with an increase of approximately 7.8824 units in popularity, holding other factors constant.

The t-value for the "num_genres" coefficient is 15.28, and the associated p-value is extremely small, indicating that the estimated coefficient is significantly different from zero. Therefore, we can conclude that the number of genres an artist is associated with has a significant impact on their popularity.

The multiple R-squared value of 0.403 indicates that the number of genres explains approximately 40.3% of the total variation in artist popularity. The adjusted R-squared value takes into account the number of predictors in the model and provides a slightly lower value of 0.4013.

The F-statistic of 233.6 with a very low p-value suggests that the overall model is statistically significant, indicating that the number of genres is a strong predictor of artist popularity.

The residuals' analysis reveals that the minimum and maximum residuals are -35.328 and 35.790, respectively, with the majority of residuals clustered around 0. This indicates that the model's assumptions regarding the distribution of residuals are reasonably satisfied.

In summary, based on this linear regression analysis, we can conclude that the number of genres an artist is associated with has a statistically significant positive effect on their popularity. However, it's important to note that the regression analysis can only show associations and not establish causality.

```{r}
model3 <- lm(popularity ~ betweenness, data=artists)

# Print model summary
summary(model3)
```
The regression model suggests that there is a statistically significant relationship between an artist's popularity and their betweenness centrality. Betweenness centrality is a measure of an artist's importance or influence in the network of collaborations. The estimated coefficient for the "betweenness" variable is 4.268e-03, indicating that, on average, a unit increase in betweenness is associated with an increase of approximately 0.004268 in popularity, holding other factors constant.

The t-value for the "betweenness" coefficient is 4.923, and the associated p-value is very small (1.32e-06). This indicates that the estimated coefficient is significantly different from zero, suggesting that betweenness centrality has a statistically significant impact on artist popularity.

The multiple R-squared value of 0.06546 indicates that the betweenness centrality explains only a small proportion (6.546%) of the total variation in artist popularity. The adjusted R-squared value of 0.06276 takes into account the number of predictors in the model and provides a slightly lower value.

The F-statistic of 24.23 with a very low p-value (1.323e-06) suggests that the overall model is statistically significant, indicating that the betweenness centrality is a statistically significant predictor of artist popularity.

The analysis of residuals shows that the minimum and maximum residuals are -46.71 and 41.12, respectively. The majority of residuals are clustered around 0, suggesting that the model's assumptions regarding the distribution of residuals are reasonably met.

In summary, based on this linear regression analysis, we can conclude that an artist's betweenness centrality has a statistically significant positive effect on their popularity. However, it's important to note that the regression analysis alone cannot establish causality, and other factors may also contribute to artist popularity.

```{r}
#Create scatterplot with regression line
ggplot(artists, aes(x=betweenness, y=popularity)) +
geom_point() +
geom_smooth(method="lm", se=FALSE)
```


```{r}
model4 <- lm(popularity ~. - artist, data=artists)

# Print model summary
summary(model4)
```
The provided output is from a linear regression analysis, where the dependent variable is "popularity" and several predictor variables are included. Here's a comment on the output:

The regression analysis suggests that the model has some predictive power in explaining the variability in artist popularity. The adjusted R-squared value of 0.4396 indicates that approximately 43.96% of the variation in popularity can be accounted for by the included predictor variables.

Among the included variables, "num_genres" shows a strong positive relationship with popularity, as evidenced by the significant coefficient estimate of 7.068. This suggests that, on average, for each additional genre an artist is associated with, their popularity is expected to increase by approximately 7.068 units.

The variable "num_collab" also has a positive relationship with popularity, although its effect is relatively smaller, with a coefficient estimate of 0.3024. Each additional collaboration an artist has is associated with a modest increase in popularity.

On the other hand, the variables "closeness" and "betweenness" do not appear to have a statistically significant effect on artist popularity. Their coefficients are not significantly different from zero, as indicated by the high p-values.

It's important to note that the coefficient estimates are subject to uncertainty, as reflected by the standard errors. The residuals, which represent the differences between the observed popularity values and the predicted values from the model, show a relatively normal distribution with a mean close to zero.

In summary, this linear regression model suggests that the number of genres an artist is associated with and the number of collaborations they have are significant predictors of popularity. However, the variables "closeness" and "betweenness" do not seem to have a significant impact on artist popularity in this model.


```{r}
your_dataframe <- artists[-c(1)]
# Step 1: Calculate Correlation Matrix
correlation_matrix <- cor(your_dataframe)

# Step 2: Visualize Correlations
heatmap(correlation_matrix, 
        cmap = colorRampPalette(c("blue", "white", "red")), 
        main = "Correlation Heatmap")

# Step 3: Calculate VIF
library(car)
vif_values <- vif(model4)  # your_lm_model represents the linear regression model

# Step 4: Calculate Tolerance
tolerance_values <- 1 / vif_values

# Step 5: Calculate Eigenvalues
eigenvalues <- eigen(correlation_matrix)$values

# Step 6: Calculate Condition Number
condition_number <- sqrt(max(eigenvalues) / min(eigenvalues))

# Step 7: Pairwise Scatterplots
pairs(your_dataframe)  # Creates scatterplot matrix for all variables in the dataframe

# Step 8: Analyze Variable Relationships
# You can use additional functions or techniques based on your specific analysis needs,
# such as partial correlation or regression diagnostics.

# Print or inspect the results
print(correlation_matrix)
print(vif_values)
print(tolerance_values)
print(eigenvalues)
print(condition_number)

```

```{r}
model7 <- lm(closeness ~ log(num_collab), data=artists)

# Print model summary
summary(model7)
```

The reason behind the strong logarithmic relationship between the number of collaborations and closeness centrality in your social network analysis is because of how social networks work.

Think of it this way: when an artist collaborates with a lot of other artists, they become more popular and influential within the network. They have a larger presence and are more likely to be involved in various collaborations.

Now, closeness centrality measures how quickly information or influence can spread from one artist to others in the network. It's like measuring how close an artist is to everyone else in terms of the shortest path lengths. Artists with high closeness centrality are considered important and influential within the network.

Here's where the logarithmic relationship comes in: as an artist starts collaborating with more and more artists, the increase in their closeness centrality becomes smaller and smaller. In simpler terms, artists who have fewer collaborations can make a big jump in their closeness centrality by collaborating with just a few more artists. But as they already have a lot of collaborations, adding more collaborations has a smaller impact on their closeness centrality.

This logarithmic relationship happens because of the way social networks are structured. They often have a few artists who are highly connected, while most artists have fewer connections. So, the impact of additional collaborations on closeness centrality diminishes logarithmically as the number of collaborations increases.

Understanding this relationship helps us see how collaborations and influence work in the social network. It shows that there are diminishing returns to collaborations in terms of increasing an artist's centrality. Other factors, like the type of collaborations or the overall network structure, also play a role in determining an artist's influence within the network.